---
title: |
  | Introduction to Data Science
  | R Tutorial
  | Week 2
author: |
  | Ilse Cuevas Andrade and Dr. Gah-Yi Ban
  | Imperial College Business School
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(tidyverse)
```

1. Deal with NAs
-----------

- Load the data from "uk_gdp_and_weeklyWage_na.csv".

- Locate the NAs and remove the rows with missing values. 
- One way to do it is using the built-in function `na.omit`

```{r}
#Load data 

uk_gdp_and_weeklyWage_na <- read.csv("data/uk_gdp_and_weeklyWage_na.csv")

uk_gdp_and_weeklyWage_no_NA<-na.omit(uk_gdp_and_weeklyWage_na)

```


2. Functions in R.
-----------

*Example*

Write a function compute_s_n that for any given n computes the sum Sn = 1^2 + 2^2 +3^2 + . . . n^2. Report the value of the sum when n = 10.

```{r}
# Two approaches to solve this question. 

compute_s_n <- function(n) {
  sum_squares<-0 # initialize vector
  # using for loop
  
  for(i in 1:n){
    sum_squares <- sum_squares +(i^2)
  }
  
  
  return(sum_squares)
}

result <- compute_s_n(10)
print(result)
```


```{r}
compute_s_n <- function(n) {
  # using vectorized approach
  sum_squares <- sum((1:n)^2)
  
  return(sum_squares)
}

# Compute S_n for n = 10
result <- compute_s_n(10)
print(result)
```

*Exercise 1*
-----------

- Load the data "uk_gdp_and_weeklyWage.csv" 
- Then, add two columns, "GDP_growth_rate" and "weeklyWageLevel", using `mutate` to show yearly GDP growth rate and if weekly wage level was over £500 or not. 

```{r Exercise 1}
uk_gdp_and_weeklyWage <- read.csv("data/uk_gdp_and_weeklyWage.csv")

uk_gdp_and_weeklyWage %>% 
  mutate(GDP_growth_rate = (GDP_m - lag(GDP_m)) / lag(GDP_m) * 100, 
         weeklyWageLevel = Weekly_pay>500)

```

*Exercise 2 (Challenging!)*
-----------

- From the dataframe `uk_gdp_and_weeklyWage`, extract a subset containing leap years only.

- *Hint*: 
    - you may want to use `rbind` together with loop and nested conditions
to extract the subset holding leap years. 
    - NULL value can be used to initialized an empty data frame.
But there are different ways to implement that.

```{r Exercise 2}

leap_year_subset <- NULL
for (i in 1:nrow(uk_gdp_and_weeklyWage)) {
  year_num <- uk_gdp_and_weeklyWage$Year[i]
  # check if it is a leap year
  if (year_num %% 100 == 0) {
    if (year_num %% 400 == 0) {
      leap_year_subset <- rbind(leap_year_subset, uk_gdp_and_weeklyWage[i, ])
    }
  } else if (year_num %% 4 == 0) {
    leap_year_subset <- rbind(leap_year_subset, uk_gdp_and_weeklyWage[i, ])
  } else {
    print(paste(year_num, "is not a leap year"))
  }
}
```



##The end