---
title: "Example"
output: pdf_document
date: "2024-11-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown &nbsp; &nbsp; &nbsp; document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:


# R Tutorial
## R Tutorial
### R Tutorial
#### R Tutorial

<!-- this is an example of how to format a comment using R markdown. -->


You can include a text for your clickable [link](https://www.imperial.ac.uk)



```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
