---
title: |
  | Introduction to Data Science
  | R exercises in-class
  | Week 1  
author: |
  | Dr. Gah-Yi Ban
  | Imperial College Business School

output:
  pdf_document: default
  html_document:
    df_print: paged
---

Getting Started
======================

**NOTE** This is an *R Notebook*, which combines text and code into a single document.
The 14 lines are the settings of this R Notebook document- please do not change anything there. 
The text in a Notebook can be given special formatting using something
called *Markdown*; it is not important to understand right away - we will take a closer
look later in the module.

Before starting any coding practice, please follow the slides to create a project
and copy the file handout01.Rmd into the project directory. Then open the file
handout01.Rmd inside the project using the *bottom-right* "Files" panel.

Code in an R Notebook is kept in *code chunks* that appear in grey
shaded boxes, like the one below.
```{r Intro to Code Chunks}
2 + 3
1 + 1
```

#```{trial}

#```


Three backticks ``` symbols are used to open and close the code chunk.
The parentheses after ``` such as {r Intro to Code Chunks}  gives it a name; think of this as 
the tag or ID of the code chunk. Note that each code chunk should have a unique name.

Run a code chunk by placing your cursor on a line of code inside it, and clicking
the **green arrow** on the top line of the code chunk. If you want to run one or more lines
in a code chunk, select the line(s) you want to run, and use **Run** 
with a down arrow and selecting "Run Selected Line(s)".

If everything is working as expected, you should see the number 5 displayed above, in a *white area directly under the code chunk*.

Notice the [1] that precedes the result of the operation. For now, it is enough to know that this number in square brackets indicates output - we'll talk more about it when we start working with vectors.

Arithmetic
----------

In the above code chunk, we used R as a calculator with 2 integer numbers. 
Use the code chunk below to try out the other common mathematical operators 
-  *  / with some float numbers.
```{r Mathematical Operators}
# Addition with 
3.5 + 2.4

# Subtraction
3.5 - 2.4

# Multiplication
3.5 * 2.4

# Division
3.5 / 0.5

```

In the code chunk above we introduced *comments*. A comment in R starts
with a hash `#` symbol and is ignored during execution. Comments can
exist on their own line, or can come immediately after some code.

**Exercise 1: Order of Operations**

R will follow the usual operator precedence that applies in mathematics.
As expected, we can also apply BODMAS And BIDMAS in the code. For example,
parentheses (round brackets) to force a specific order of evaluation.

In the code chunk below, *add* or *remove* parentheses so that all three expressions evaluate to the same result: 20
```{r Exercise 1}
# Fix the parentheses below
(2 + 3 + 5)*2
4*(10 - 5)
2*8 + 4
```


**NOTE:** Use **whitespace** above to improve the readability our
code for everyone (both your future selves and your tutors!) 

Best practice:

- a space between numbers and operators
- a space between operators and parentheses
- no space between numbers and parentheses

Variables
---------

Often, it is useful to store data, or the result of an operation. Like other
programming languages, R supports the *assignment* of *variables* for this.

For now, you can think of variables as having a name, and containing some data.
To assign data to a variable in R, we use the *arrow assignment operator*.
```{r Variable assignment}
# Assign a numeric value to x
x <- 3.14

# Assign the result of an operation to y
y <- 5 + 1.86

# Sum the variables and assign the result to z
z <- x + y

# Print output
print(z)

```

**NOTE:** assigning data to variables does not produce any output. In the top-
right pane of RStudio, the *Environment* window (**top-right** panel) will show any variables you
created, so take a look there after running the above code chunk. Alternatively, we can always print a variable to see its value.

Now using the idea of variables, *add* or *remove* parentheses and mathematical operators in the code chunk below so that all three expressions evaluate to the same result: 20

```{r Exercise 2}
# define variables
a <- 2
b <- 3
c <- 5
# Apply BIDMAS below to obtain 20 in each line
(a + b + c)*a
(a + a)*(c*a - c)

```

We can re-assign variables, but note that this will overwrite the original value.
```{r Re-assigning variables}
# Assign a new numeric value to z
z <- 22

# Print z
print(z)

# Assign a new character value to z
z <- "Hello!"

# Print z
print(z)

```

**NOTE:** Variables hold different *types* of data. In the above example, `z`
variable was used to first hold numeric data, then character data.


## The End

