---
title: |
  | Introduction to Data Science
  | R exercises in-class
  | Week 3  
author: |
  | Dr. Gah-Yi Ban
  | Imperial College Business School

output:
  pdf_document: default
  html_document:
    df_print: paged
---


Tidyverse continued
====================

Let us start by calling the `tidyverse` library.

**NOTE:** If you don't have the tidyverse package on your RStudio, please first install it by running `install.packages("tidyverse")` in the R Console.

```{r load lib, warning=F, message=F}
library(tidyverse)
```

```{r flight}

flights %>% arrange(desc(dep_delay))


```

Before starting any practice and exercises, load the dataset "uk_gdp_and_weeklyWage.csv". Recall this gives us data on the average weekly pay (in £) and the total GDP (in millions of £) from 2000 to 2022.


```{r load data}
uk_gdp_and_weeklyWage <- read.csv("./data/uk_gdp_and_weeklyWage.csv", 
                                  stringsAsFactors = F)

head(uk_gdp_and_weeklyWage)
```

With pipe operators, you may use one or more of the following verbs (note there are others):

  + `select()`: pick variables based on their names.
  + `filter()`: pick cases based on their values.
  + `mutate()`: add new variables that are functions of existing variables.
  + `arrange()` allows you to sort a data frame by a particular column.
  + `summarise()`: reduce multiple values down to a single summary.
  + These all combine naturally with `group_by()` which allows you to perform any
  operation “by group”. 
  
**Note ** If you use `group()`, then `ungroup()` is always used after performing all the calculations. If you forget to `ungroup()` data, future data management will likely produce errors. Always `ungroup()` swhen you’ve finished with your calculations.

### mutate

Let us add GDP growth rates and categorize the growth rate into quartiles.

*Note:* there will be NA(s) in the GDP growth rates data

To do so, we need to make use of `cut()` and `quantile()` functions.

The `cut()` function in R is used to cut a range of values into bins and specify labels for each bin. This function uses the following syntax:

**cut(x, breaks, include.lowest = TRUE or FALSE, labels = ...)**

where:

* x: Name of vector

* breaks: Number of breaks to make or vector of break points

* labels: Labels for the resulting bins. `labels = NULL` produces interval bins and `labels = FALSE` produces integer category labels starting at 1. Otherwise, you can assign your own labels with `labels = your own vector of labels`.

* include.lowest: Boolean value to include lowest break value 

To see how `cut()` works, study and run the code below that puts a vector of numbers from 0 to 100 into 5 bins.
```{r cut}
vec <- c(0:100)

cut(vec, c(0,20,40,60,80,100), include.lowest = TRUE, labels = FALSE) 

```

The `quantile()` function produces sample quantiles corresponding to the given probabilities. The smallest observation corresponds to a probability of 0 and the largest to a probability of 1.

The syntax is:

**quantile(x, probs, na.rm = TRUE or FALSE, type = 7)**

where:

* x: Numeric vector whose sample quantiles are wanted. NA and NaN values are not allowed in numeric vectors unless na.rm is set to TRUE.

* probs: Numeric vector of probabilities. E.g., if `probs = seq(0, 1, 0.25)`, then you are asking for the quartiles (0%, 25%, 50%, 75% and 100%).  

* na.rm = TRUE or FALSE: if TRUE, any NA and NaN's are removed from x before the quantiles are computed.

* type: An integer between 1 and 9 selecting one of the nine quantile algorithms available are to be used. Default is to keep it at 7.

To see how `quantile()` works, study and run the code below that puts a vector of numbers from 0 to 100 into quartiles.
```{r quantile}

vec <- c(0:100)

quantile(vec, probs = seq(0, 1, 0.25), type = 7) 

```

Now we are ready to use `mutate` to add GDP growth rates and categorize the growth rate into quartiles. Study and run the code chunk below. 

```{r mutate}

uk_gdp_and_weeklyWage <- uk_gdp_and_weeklyWage %>% 
  mutate(GDP_growth_rate = (GDP_m - lag(GDP_m)) / lag(GDP_m) * 100, 
         )

uk_gdp_and_weeklyWage <- uk_gdp_and_weeklyWage %>% 
  mutate(GDP_growth_rate_Quartile = cut(GDP_growth_rate, 
                                  quantile(GDP_growth_rate, 
                                           probs = seq(0, 1, 0.25), type = 7, 
                                           na.rm = T), 
                                  include.lowest = T, labels = F)
          )

head(uk_gdp_and_weeklyWage)

```


### arrange

Let's sort the data by the order of GDP growth rates using arrange.

```{r sort data}
uk_gdp_and_weeklyWage %>% arrange(GDP_growth_rate) %>% head()

```

### group_by

Let us now investigate the mean, standard deviation of GDPs, mean and standard deviation of weekly wage, and mean and standard deviation of GDP growth rates 
in each GDP growth rate quartile group by `summarise`.

Can you do this in the following code chunk?

*Exercise 1*

```{r Exercise 1}
uk_gdp_and_weeklyWage %>% 
  ???(???) %>%
  ???(mean_GDP = ???, std_GDP = ???,
            mean_wage = ???, std_wage = ???,
            mean_GDPrate = ???, 
            std_GDPrate = ???) %>%
  ungroup()
```

After running the above code chunk, have you observed any issue(s)? 

To address the issue(s), adjustment in the above code chunk can be done by different ways. For example, `filter` can be added to implement that in the Exercise 2.

*Exercise 2*
```{r Exercise 2}
uk_gdp_and_weeklyWage %>% 
  filter(???) %>%
  ???(???) %>%
  ???(mean_GDP = ???, std_GDP = ???,
            mean_wage = ???, std_wage = ???,
            mean_GDPrate = ???, 
            std_GDPrate = ???) %>%
  ungroup()
```

Sometimes, viewing the first or last few rows of the data will be helpful to 
understand the data. Recall that we used `head` to display the first few rows 
of data.But we often need to sort the data before viewing. 

Let us try viewing the first 10 rows of "uk_gdp_and_weeklyWage" after sorting 
the "GDP_growth_rate" with `arrange` in Exercise 3.

*Exercise 3*
```{r Exercise 3}
uk_gdp_and_weeklyWage %>% ??? %>% ???
```
Now let us try viewing the last 10 rows of "uk_gdp_and_weeklyWage" after sorting the "GDP_growth_rate" with `arrange` and `tail` in Exercise 4.

*Exercise 4*
```{r Exercise 4}
uk_gdp_and_weeklyWage %>% ??? %>% ???
```

### summarise

`summarise()` or `summarize()` function provides a summary of data, such as
mean, standard deviation, median etc. The function is often used together with 
`group_by` if the data contain groups.

Let's start with a simply task by showing the mean of weekly wage, GDP and GDP growth rate.

```{r summarize}
uk_gdp_and_weeklyWage %>% 
    summarise(mean_wage = mean(Weekly_pay), mean_GDP = mean(GDP_m), 
              mean_GDP_growth = mean(GDP_growth_rate))
```
Have you found any issue after running the code above? Why? What's the solution?

You can handle NA values by adding an argument `na.rm=TRUE` or `na.rm=T` (for shorter version) into the function `mean`. Try these below.

*Exercise 5*

```{r Exercise 5}
uk_gdp_and_weeklyWage %>% 
    summarise(mean_wage = mean(Weekly_pay), mean_GDP = mean(GDP_m), 
              mean_GDP_growth = ???)

```

If we run `summary()` on the GDP growth rate column, we get 2.163% as the median
rate. Let us extract this number from the string output using the `parse_number()` function, as below. To find out how to use `parse_number()`, you can search online or type `?parse_number` in the Console.

```{r GDP growth rate median}
GDP_growth_rate_summary <- uk_gdp_and_weeklyWage %>% 
  summary(GDP_growth_rate, na.rm=T)

GDP_growth_rate_summary

GDP_growth_rate_median <- parse_number(GDP_growth_rate_summary[???,???], 
                                       locale = locale(grouping_mark = " "))

GDP_growth_rate_median
```

Let's group the GDP growth rate data by the benchmark at 2.163% and check the mean and standard deviation of the GDP growth rate.

```{r group summary}
uk_gdp_and_weeklyWage <- uk_gdp_and_weeklyWage %>%
  mutate(lower_median = GDP_growth_rate < GDP_growth_rate_median)

uk_gdp_and_weeklyWage %>% group_by(lower_median) %>%
  summarise(mean_GDP_growth = mean(GDP_growth_rate, na.rm=T), 
            std_GDP_growth = sd(GDP_growth_rate, na.rm = T)) %>%
  ungroup()
```

Any issue(s) found from the code chunk above?

Please find a solution to address the issue in the following code chunk.

*Exercise 6*

```{r Exercise 6}
uk_gdp_and_weeklyWage <- uk_gdp_and_weeklyWage %>%
  mutate(lower_median = GDP_growth_rate < GDP_growth_rate_median)

uk_gdp_and_weeklyWage %>% 
  group_by(lower_median) %>%
  ??? %>%
  summarise(mean_GDP_growth = mean(GDP_growth_rate, ???), 
            std_GDP_growth = sd(GDP_growth_rate, ???)) %>%
  ungroup()
```

## The End
