---
title: |
  | Introduction to Data Science
  | R Tutorial
  | Week 3
author: |
  | Ilse Cuevas Andrade and Dr. Gah-Yi Ban
  | Imperial College Business School

output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
library(dplyr)
```


*Exercise* 

Load built-in EuStockMarkets dataset, which contains daily closing prices of major European stock indices. The goal is to practice using the summarise(), group_by(), and arrange() functions to analyse financial data.

###Dataset:
The EuStockMarkets dataset contains data on four major European stock indices from 1991 to 1998:

**DAX**: German stock index.
**SMI**: Swiss stock index.
**CAC**: French stock index.
**FTSE**: British stock index.


```{r}
#Load the data
data("EuStockMarkets")
# Transform to data frame
stock_data <- as.data.frame(EuStockMarkets)
# Data processing. Adding the date - time series data.
stock_data$Date <- seq(as.Date("1991-01-01"), by = "days", length.out = nrow(stock_data))
#Print the first 6 rows
head(stock_data)
```

*Exercise 1*
Extract the year from the Date column and group the data by Year. Then, calculate the average closing price of the four indices for each year.

```{r Exercise 1}


```
*Exercise 2*
Find the year with the highest average DAX index and sort the years in descending order based on the average DAX value.

```{r Exercise 2}

```
*Exercise 3*
In this exercise, you will work with the **`LifeCycleSavings`** dataset and use functions like `pull()`, the dot operator (`.`), `case_when()`, and `between()` to analyze economic data. This dataset contains economic information on savings rates in different countries.

### Dataset
The `LifeCycleSavings` dataset contains the following variables:
- **sr**: Aggregate personal savings as a percentage of disposable income.
- **pop15**: Percentage of population under 15 years of age.
- **pop75**: Percentage of population over 75 years of age.
- **dpi**: Real per-capita disposable income.
- **ddpi**: Percentage growth rate of dpi.

You can load and convert the dataset to a data frame with the following command:

```{r Exercise 3}
data("LifeCycleSavings")
savings_data <- as.data.frame(LifeCycleSavings)
```

Create a new column savings_category to classify countries based on their savings rate (sr) as follows:

"Low Savings" if sr is less than 10.
"Moderate Savings" if sr is between 10 and 20.
"High Savings" if sr is greater than 20.

```{r Exercise 4}

```
*Example*
```{r}
# Extract the 'sr' column using pull()
savings_rate <- savings_data %>%
  pull(sr)

# Display the first 10 values
head(savings_rate, 10)

```


```{r}
# Access the 'sr' column using the dot operator
savings_rate <- savings_data %>%
  .$sr

# Display the first 10 values
head(savings_rate, 10)
```

## The end
